The fraction of polar aligned circumbinary disks

### 2025-04-07 src/tex/response.tex

\documentclass{letter}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath}
\usepackage{bm}

\fboxrule=3pt
\fboxsep=4pt

\newenvironment{response}[1]{
    \fcolorbox{blue}{white}{
    \begin{minipage}{0.46\textwidth}
        #1
    \end{minipage}
    }\hfill
    \begin{minipage}{0.46\textwidth}
    }
    {\end{minipage}}

\begin{document}

\begin{letter}{}
    \opening{Dear Editor Carraro, Data Editors, and Anonymous Reviewer,}

    Please find our revised manuscript alongside the below response to the reviewers report. The authors would like to thank the editor, reviewer, and data editors for their thoughtful review, which has lead to a greatly improved manuscript. Changes have been highlighted in \textcolor{blue}{\bf blue}.

    \begin{response}{Reviewer's Comment}
        Authors' Response
    \end{response}

    \begin{response}
        {
            L35-38: The evidence pointed out by the authors for AC Her is extremely indirect and also around an evolved binary star with a CBD, for which the disc formation pathway and subsequent evolution remains very unclear. At the moment, this "evidence" is loosely supported by data as the planet was not confirmed/observed but simply proposed as the "most likely scenario" (Martin et al., 2023) based on disc truncation arguments alone. Therefore, I would strongly suggest to rephrase this part to avoid conveying the wrong idea to non-specialists.
        }
        We have changed the order of the text and changed the language so that this portion is clearer.
    \end{response}

    \begin{response}
        {L93: A bit too self-referential in my opinion. SPH simulations on this very topic were also done by other teams (Aly+2015; Cuello+2019, Young+2023, Ceppi+2023). It is not necessary to be exhaustive here, but I would avoid citing their own work exclusively.}
        Thank you for these suggestions. We have added the references.
    \end{response}

    \begin{response}
        {L128: "Code implementing" typo? If so, please rephrase.}
        This is not a typo, but has been rephrased for clarity.
    \end{response}

    \begin{response}
        {L142: Equation 2, I believe two "if" statements are missing.}
        These statements have been added.
    \end{response}

    \begin{response}
        {L184: Perhaps specify that "isotropic angular momentum distribution" means that the relative angle between the disc and binary orbital planes is uniformly distributed. This would help to give a more clear astrophysical interpretation.}
        This description is not entirely correct. By ``isotropic'' we mean that the specific angular momentum vector $\bm{\ell}$ has no preference towards any infinitesimal solid angle $d\Omega$. We have added text to clarify.
    \end{response}

    \begin{response}
        {L239: I would suggest defining or explaining how quadrants are displayed. I assume the anti-clockwise convention is used here.}
        This is correct. We have added some text to specify this convention.
    \end{response}

    \begin{response}
        {Fig. 6: Please comment the insert.}
        We have added a more descriptive caption in addition to a description of the insert.
    \end{response}

    \begin{response}
        {Fig. 7: the caption could be more descriptive to guide the reader.}
        We have added a more descriptive caption.
    \end{response}

    \begin{response}
        {L426: drops instead of drop?}
        This has been fixed. Thank you for catching it.
    \end{response}

    \begin{response}
        {L445: In the definition of f, is M meant to be M\_b or M\_1 + M\_2? It is likely the latter one. Unless I am mistaken, it was not defined earlier.}
        It is supposed to be $M_\text{b}\equiv M_1 + M_2$. We have added text for clarification.
    \end{response}

    \begin{response}
        {L468: Again, I think m\_r was not defined earlier. I assume this is the mass of the ring. Also, would it be possible to write down the expression of Gamma\_2 to effectively see that it tends to 1 when j is large? Since it is very complicated it could go in the Appendix to avoid interrupting the flow of the section.}
        We have added text to define $m_\text{r}$. We also added a section in the appendix to describe the process of computing $\Gamma_2$. However, we also clarified in the text why $\Gamma_2$ tends to 1 for small $j$, so this section may no longer be necessary.
    \end{response}

    \begin{response}
        {L473: Syntax/repetition issue. Please rephrase.}
        We have fixed the issue by clarifying that we are translating the precession timescale from code units to physical units.
    \end{response}

    \begin{response}
        {Fig. 8: The bottom part of this figure could be further commented in Sect. 6.1.}
        We have added text in Section 6.1 discussing this panel.
    \end{response}

    \begin{response}{
        L532: Bohn+2022 could be a relevant reference to add here.
    }
     Thank you for the great suggestion. We have added the reference.   
    \end{response}

    \begin{response}
        {L580: I would conclude the discussion (or add to the conclusion) that, besides imposing constraints on the disc radial extend and j, the two studied mechanisms (disc breaking and self-gravity) are expected to lower the fraction of polar CBDs. If I understood correctly, large values of j and radially extended discs are unlikely to become polar (as a whole at least). Therefore, the statistical relevance/weight of this subgroup of discs should be less important at the overall polar CBD population level. This could be briefly commented at the end.}
        This is a great suggestion, however I think the overall effect is a bit different. High-$j$ disks have polar probabilities that are sensitive to initial inclination and insensitive to $e_{\rm b}$. Radially extended disks have long alignment timescales and may not evolve much within their lifetimes. Reducing these two populations (which should have significant overlap) would make $f_\text{polar}$ more sensitive to $e_{\rm b}$ and reduce the population of disks that end their lives outside of polar or coplanar alignments.
    \end{response}

    \begin{response}{
        L615: Unclear what "1/s" means. Below a second per calculation/timestep? Please specify.
    }
        We have clarified that this refers to the time to run a single simulation until it can be determined if the third body is undergoing circular precession or libration.
    \end{response}

    \begin{response}{
        L628: Sections (missing s).
    }
    This has been fixed.
    \end{response}

    \begin{response}{
        L642: At the end of the conclusions (or else at the end of the discussion), a brief discussion on how future observations could test these findings would add value to the article.
    }
        We have added a discussion about the possibility of confirming this work though observations. We propose a survey to measure the inclinations of disks in young eclipsing binary systems could reasonably constrain the polar fraction.
    \end{response}

\end{letter}

\end{document}

### 2025-04-14 src/tex/response.tex

    \begin{response}
        {L93: A bit too self-referential in my opinion. SPH simulations on this very topic were also done by other teams (Aly+2015; Cuello+2019, Young+2023, Ceppi+2023). It is not necessary to be exhaustive here, but I would avoid citing their own work exclusively.}
        Thank you for these suggestions. We have added the references.
    \end{response}

    \begin{response}
        {L128: "Code implementing" typo? If so, please rephrase.}
        This is not a typo, but has been rephrased for clarity.
    \end{response}

    \begin{response}
        {L142: Equation 2, I believe two "if" statements are missing.}
        These statements have been added.
    \end{response}

    \begin{response}
        {L184: Perhaps specify that "isotropic angular momentum distribution" means that the relative angle between the disc and binary orbital planes is uniformly distributed. This would help to give a more clear astrophysical interpretation.}
        By ``isotropic'' we mean that the specific angular momentum vector $\bm{\ell}$ has no preference towards any infinitesimal solid angle $d\Omega$. We have added text to clarify this.
    \end{response}

@@

    \begin{response}
        {L473: Syntax/repetition issue. Please rephrase.}
        We have fixed the issue by clarifying that we are translating the precession timescale from code units to physical units.
    \end{response}

    \begin{response}
        {Fig. 8: The bottom part of this figure could be further commented in Sect. 6.1.}
        We have added text in Section 6.1 discussing this panel.
    \end{response}

    \begin{response}{
        L532: Bohn+2022 could be a relevant reference to add here.
    }
     Thank you for the great suggestion. We have added the reference.   
    \end{response}

    \begin{response}
        {L580: I would conclude the discussion (or add to the conclusion) that, besides imposing constraints on the disc radial extend and j, the two studied mechanisms (disc breaking and self-gravity) are expected to lower the fraction of polar CBDs. If I understood correctly, large values of j and radially extended discs are unlikely to become polar (as a whole at least). Therefore, the statistical relevance/weight of this subgroup of discs should be less important at the overall polar CBD population level. This could be briefly commented at the end.}
        This is a great suggestion, however I think the overall effect is a bit different. High-$j$ disks have polar probabilities that are sensitive to the initial inclination and insensitive to $e_{\rm b}$. Radially extended disks have long alignment timescales and may not evolve much within their lifetimes. Reducing these two populations (which should have significant overlap) would make $f_\text{polar}$ more sensitive to $e_{\rm b}$ and reduce the population of disks that end their lives outside of polar or coplanar alignments. We have added some discussion of this to the paper.
    \end{response}

### 2025-04-14 src/tex/ms.tex
Rebecca's edits on reviewed version

Observed circumbinary planets -- those orbiting two stars -- have mostly been found to orbit in the same plane as their hosts \citep{doyle2011,orosz2012,welsh2012}. While this preference is the result of target selection effects (\citealt{martin2014}, see also \citealt{martin2015}, \citealt{martin2017a}), it is also representative of the fact that, for the disk progenitors of these planets \citep[from which the planets inherit orbital parameters, see][]{childs2021}, a coplanar alignment is one of two possible stable orientations \citep{bate2000,lubow2000}. The other configuration -- a polar disk -- is characterized by being inclined $90^\circ$ relative to the binary orbital plane \citep{aly2015,martin2017}. \RR{O}nly a handful of confirmed polar disks have ever been discovered. These include a polar debris disk in the 99 Her system (\citealt{kennedy2012}, which likely evolved from a polar gas disk, see \citealt{smallwood2020}) and polar gas disks in the HD 98800 \citep{kennedy2019} and V773 Tau \citep{kenworthy2022} systems. \RR{Evidence for one polar planet has been found via indirect means \citep[AC Her, ][]{hillen2015,anugu2023,martin2023}}

@@

\begin{figure*}[!htbp]
    \begin{centering}
        \includegraphics[width=1\textwidth]{figures/compare_ebin.pdf}
        \caption{
            Polar fraction as a function of $e_{\rm b}$ for various values of $j$. Each line is computed via a numerical integration of equation (\ref{eq:prob_polar}), assuming an isotropic distribution of $i$.
            In the low-$j$ regime the polar fraction is a strong function of $e_{\rm b}$. However, as $j$ increases, $f_{\rm polar}$ becomes insensitive to the binary eccentricity (except for the case that $e_\text{b}\approx 1$). Note that $f_\text{polar}(e_\text{b}\approx 0)$ is maximized when $j=0.5$.
        }
        \script{compare_ebin.py}
        \label{fig:ebin}
    \end{centering}
\end{figure*}

@@

\begin{figure*}
    \begin{centering}
        \includegraphics[width=\textwidth]{figures/compare_ebin_idist.pdf}
        \caption{
            Same as Figure \ref{fig:ebin}, but with the mutual inclination distribution given by equation (\ref{eq:idist}). \RR{When $j$ is low, lower inclination preferences decrease $f_\text{polar}$ around $e_\text{b}\sim0.5$, but have little effect near 0 and 1. In the high-$j$ regime, low inclinations lower the value that $f_\text{polar}$ tends to as $e_\text{b}\rightarrow 0$. {\bf Insert}: Mutual inclination probability distribution normalized by $\sin i$. This normalization shows that the probability-per-unit-area on the unit sphere becomes very high at $i=0$ (coplanar alignment) when $\sigma$ is small (dashed and dotted lines), but it is uniform everywhere in the isotropic case ($\sigma \rightarrow \infty$, solid line).}
        }
        \script{compare_ebin_idist.py}
        \label{fig:ebin-idist}
    \end{centering}
\end{figure*}

@@

Note that $\Gamma_1$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle. For $j>0$, we must compute the precession timescale ourselves. The time coordinate used in Section \ref{sec:rk} is $\tau \equiv \alpha' t$. $\alpha'$ is defined by \citet[see eq. 3.9]{farago2010}, and we write it as
\begin{equation}
    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \,\Gamma_2(j,f,a_\text{b}/r, e_\text{b}) \, ,
\end{equation}
where we have written the $\sqrt{1+m_\text{r}/M_\text{b}}$ factor as $\Gamma_2$. \RR{$m_\text{r}$ is the mass of the ring. To compute $\Gamma_2$ from our variables requires finding the roots of a cubic equation (see equation (\ref{eq:gamma2-eq})). However, for low $j$ (e.g. $<10$), $m_\text{r} \ll M_\text{b}$ and $\Gamma_2 \sim 1$.}

>>>>

Observed circumbinary planets -- those orbiting two stars -- have mostly been found to orbit in the same plane as their hosts \citep{doyle2011,orosz2012,welsh2012}. While this preference is the result of target selection effects (\citealt{martin2014}, see also \citealt{martin2015}, \citealt{martin2017a}), it is also representative of the fact that, for the disk progenitors of these planets \citep[from which the planets inherit orbital parameters, see][]{childs2021}, a coplanar alignment is one of two possible stable orientations \citep{bate2000,lubow2000}. The other configuration -- a polar disk -- is characterized by being inclined $90^\circ$ relative to the binary orbital plane \citep{aly2015,martin2017}. \RR{O}nly a handful of confirmed polar disks have ever been discovered. These include a polar debris disk in the 99 Her system (\citealt{kennedy2012}, which likely evolved from a polar gas disk, see \citealt{smallwood2020}) and polar gas disks in the HD 98800 \citep{kennedy2019} and V773 Tau \citep{kenworthy2022} systems. \RR{Currently the only evidence for a polar circumbinary planet comes from AC Her, a post-AGB star system that hosts an unusually-truncated circumbinary disk, with a planet located interior to the disk's inner edge being the most likely cause \citep{hillen2015,anugu2023,martin2023}.}

@@

\begin{figure*}[!htbp]
    \begin{centering}
        \includegraphics[width=1\textwidth]{figures/compare_ebin.pdf}
        \caption{
            Polar fraction as a function of $e_{\rm b}$ for various values of $j$. Each line is computed via a numerical integration of equation (\ref{eq:prob_polar}), assuming an isotropic distribution of $i$.
       For low-$j$, the polar fraction is a strong function of $e_{\rm b}$. However, as $j$ increases, $f_{\rm polar}$ becomes insensitive to the binary eccentricity (except for the case that $e_\text{b}\approx 1$). Note that $f_\text{polar}(e_\text{b}\approx 0)$ is maximized when $j=0.5$.
        }
        \script{compare_ebin.py}
        \label{fig:ebin}
    \end{centering}
\end{figure*}

@@

\begin{figure*}
    \begin{centering}
        \includegraphics[width=\textwidth]{figures/compare_ebin_idist.pdf}
        \caption{
            Same as Figure \ref{fig:ebin}, but with the mutual inclination distribution given by equation (\ref{eq:idist}). \RR{  The insert shows the mutual inclination probability distribution normalized by $\sin i$. This normalization shows that the probability-per-unit-area on the unit sphere becomes very high at $i=0$ (coplanar alignment) when $\sigma$ is small (dashed and dotted lines), but it is uniform everywhere in the isotropic case ($\sigma \rightarrow \infty$, solid line).   For low $j$, lower inclination preferences decrease $f_\text{polar}$ around $e_\text{b}\sim0.5$, but have little effect near $e_{\rm b}=0$ and 1. For high-$j$, low inclinations lower the value that $f_\text{polar}$ tends to as $e_\text{b}\rightarrow 0$.}
        }
        \script{compare_ebin_idist.py}
        \label{fig:ebin-idist}
    \end{centering}
\end{figure*}

@@

Note that $\Gamma_1$ depends on $e_\text{b}$, and so it is only a constant in the case of a test particle. For $j>0$, we must compute the precession timescale ourselves. The time coordinate used in Section \ref{sec:rk} is $\tau \equiv \alpha' t$. $\alpha'$ is defined by \citet[see eq. 3.9]{farago2010}, and we write it as
\begin{equation}
    \alpha' \equiv \frac{3}{4} \Omega_\text{b} \left(\frac{a_\text{b}}{r}\right)^{7/2} f(1-f) \,\Gamma_2(j,f,a_\text{b}/r, e_\text{b}) \, ,
\end{equation}
where we have written the $\sqrt{1+m_\text{r}/M_\text{b}}$ factor as $\Gamma_2$. \RR{$m_\text{r}$ is the mass of the ring. To compute $\Gamma_2$ from our variables requires finding the roots of a cubic equation (see equation (\ref{eq:gamma2-eq})). However, for low $j$ (e.g. $j<10$), $m_\text{r} \ll M_\text{b}$ and $\Gamma_2 \sim 1$.}
